---
title: "Sample analysis"
output: html_notebook
---

# Packages

```{r}
library(tidyverse)
library(sf)
library(plotly)
library(leaflet)
```

## Functions

```{r}
`%<>%` <- magrittr::`%<>%`
```

## API tokens

```{r}
mapbox <- "https://api.mapbox.com/styles/v1/robertmitchellv/cipr7teic001aekm72dnempan/tiles/256/{z}/{x}/{y}@2x?access_token=pk.eyJ1Ijoicm9iZXJ0bWl0Y2hlbGx2IiwiYSI6ImNpcHI2cXFnbTA3MHRmbG5jNWJzMzJtaDQifQ.vtvgLokcc_EJgnWVPL4vXw"
```

# Data

## Read the affordable housing data

```{r}
housing <- read_csv("../data/raw/Affordable_Rental_Housing_Developments.csv") %>%
  janitor::clean_names()
```

__Check for problems__
```{r}
problems(housing)
```

## Read the geo data

```{r}
geo_data <- read_sf(
  dsn = "../data/geo/chicago_boundaries_community_areas.geojson",
  layer = "chicago_boundaries_community_areas"
)
```

# Tidy

## Get summary totals

* Before joining data, we need totals for the map, e.g., how many veterans projects in a particular community?
  + There are 24 different property types; most of it is senior and family, maybe we can simplify?
  + There are 146 different developpers with the top 17 all developping and managing 5+ affordable housing properties

### Look for categories

```{r}
housing %>% 
  count(property_type) 
```

> This is subjective, but it is also important to remember that _all_ classification or categorization is subjective--it's a critical choice and important to remember when working with data. Being clear or upfront about this kind of work, I feel, is good. I will even comment code like this with why I'm doing it this way in case I need to update or rework later.

```{r}
# We'll use these list with dplyr to assign a new variable
artist <- c(
  "Artist Housing", "Artist Live/Work Space"
)

disabled <- c(
  "People with Disabilities" 
)

family <- c(
  "Multfamily", "Multifamily", "Multifamily/Artists", "Multigfamily",
  "Mutifamily", "Inter-generational" 
)

senior <- c(
  "Senior", "Senior HUD 202", "Senior LGBTQ", "Seniors"
)

supportive <- c(
  "65+/Supportive", "Disabled/Homeless", "Supportive", "Supportive Housing",
  "Supportive/HIV/AIDS", "Supportive/Kinship Families", "Supportive/Males 18-24yrs.",
  "Supportive/Teenage Moms", "Supportive/Veterans", "Supportive/Youth/Kinship Families"
)
```

__Using `dplyr` we will check if the `property_type` is in the named list; if it is, then we'll assign the string _after_ the `~` to the variable--I usually also try to catch errors by assigning anything left over to `"error"` to see what went wrong__
```{r}
housing %<>%
  mutate(property_category = case_when(
    property_type %in% artist ~ "Artist",
    property_type %in% disabled ~ "Disabled",
    property_type %in% family ~ "Family",
    property_type %in% senior ~ "Senior",
    property_type %in% supportive ~ "Supportive",
    property_type == "ARO" ~ "ARO", # what is ARO??
    TRUE ~ "error"
  )) 
```

### Totals for `property_category`

```{r}
housing %<>%
  group_by(property_category) %>%
  mutate(n_by_property_category = n())
```

### Totals for `management_company` for each community

* Group by the company and by the community
* Create a variable with the count of that grouping
* Organize by community name
* Remove the grouping
* Group by the community only
* Create a variable with the count of that grouping
* Remove the grouping

```{r}
housing %<>% 
  group_by(management_company, community_area_name) %>% 
  mutate(n_by_management_company = n()) %>%
  arrange(community_area_name) %>%
  dplyr::ungroup() %>%
  group_by(community_area_name) %>%
  mutate(n_affordable_housing = n()) %>%
  dplyr::ungroup()
```

## Join

* Join data (on community area name)
  + `geo_data` all caps; use `stringr::str_to_title()` to correct
  + `housing` variable name too long; shorte to `community` to match `geo_data`

```{r}
geo_data %<>%
  mutate(community = str_to_title(community))
```

```{r}
housing %<>%
  rename(community = community_area_name)
```

__Join and make sure the data returned still has the simple features class attribute__
```{r}
data <- left_join(geo_data, housing, by = "community")
"sf" %in% class(data)
```

# Create maps

```{r}
icons <- awesomeIcons(
  icon = "fa-home",
  library = "fa",
  markerColor = "orange",
  iconColor = "white"
)

leaflet(data) %>%
  addTiles(mapbox) %>%
  addPolygons(
    fillColor = "grey",
    weigh = 1.5,
    fillOpacity = 0.7,
    smoothFactor = 0.5,
    color = "white",
    label = ~community) %>%
  addAwesomeMarkers(
    icon = icons,
    lng = ~data$longitude, lat = ~data$latitude)
  
```

